# Week 1 Data manipulation and EDA {#eda}

## Introduction

```{r, include = F}
library(kableExtra)
options(kableExtra.html.bsTable = T)
```


```{r, message = FALSE, warning = FALSE}
library(ISLR)
library(tidyverse)
library(haven)
library(readxl)
```

## Data types

### Exercise 1

```{r}
object_1 <- 1:5
object_2 <- 1L:5L
object_3 <- "-123.456"
object_4 <- as.numeric(object_2)
object_5 <- letters[object_1]
object_6 <- as.factor(rep(object_5, 2))
object_7 <- c(1, 2, 3, "4", "5", "6")
```

### Exercise 2

```{r}
as.numeric(object_7)
```

## Lists and data frames

### Exercise 3

```{r}
objects <- mget(ls(pattern = "^object_\\d{1}$"), .GlobalEnv)
```

### Exercise 4

```{r}
df <- data.frame(object_1, object_2, object_5)
```

### Exercise 5

```{r}
ncol(df)
nrow(df)
```
## Loading, viewing and summarising data

### Exercise 6

```{r, message = F}
apps <- read_csv("data/googleplaystore.csv")
```

### Exercise 7

Yes, the variables `size`, `installs` and `price` could have been numeric variables, rather than character variables. 

### Exercise 8

```{r}
knitr::kable(head(apps[,-1]), booktabs = T) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```


### Exercise 9

```{r}
students <- readxl::read_xlsx("data/students.xlsx")
students$student_number <- as.character(students$student_number)

head(students) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
### Exercise 10

```{r}
summary(students)
```

## Data transformation with `dplyr`

### Exercise 11

```{r}
students %>%
  filter(grade < 5.5)
```

### Exercise 12

```{r}
students %>%
  filter(grade > 8 & programme == "A")
```

### Exercise 13

```{r}
students %>%
  arrange(programme, desc(grade))
```

### Exercise 14

```{r}
students %>%
  select(student_number, programme)
```

### Exercise 15

```{r}
students_recoded <- students %>%
  mutate(programme = recode(programme, "A" = "Science", "B" = "Social Science"))
```

### Exercise 16

```{r, message = F}
popular_apps <- read_csv("data/googleplaystore.csv") %>%
  mutate(Installs = parse_number(Installs)) %>%
  filter(Installs > 500000000) %>%
  arrange(desc(Rating)) %>%
  select(Category, Rating, Reviews, Size, Installs, Type, Price)
```

### Exercise 17

```{r}
popular_apps %>%
  summarise_if(is.numeric, 
               .funs = list(median = median, min = min, max = max))
```

### Exercise 18

```{r}
mad <- function(x) median(abs(x - median(x)))

popular_apps %>%
  summarise_if(is.numeric, 
               .funs = list(median = median, 
                            min = min, 
                            max = max,
                            mad = mad))
```

### Exercise 19

```{r}
popular_apps %>%
  group_by(Category) %>%
  summarise(mean = mean(Rating),
            median = median(Rating),
            sd = sd(Rating),
            min = min(Rating),
            max = max(Rating),
            mad = mad(Rating))
```

